# Introduction
The data are about ‘Smart Start’ Program  in the university to review entry level courses (Physics, Biology (Human Anatomy), and Spanish). 

First, I review research questions, import and clean data, as well as describe the data and variables.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment='')
```

```{r warning=FALSE, message=FALSE}
rm(list =ls())
library(tidyverse)
```


# Research questions

1.What are our overall success rates (C or better)?

2.What student information helps predict success in their courses?

3.Are success rates improving or getting worse?

4.Are some of their instructors better than others? 

5.Does teaching frequently the course improve instructor performance over time?

6.How do their students, in their academic plans, perform compared to students from other departments/colleges?


# Description

## Data

It includes five tables of data: acad_plan, class_instructors, class_inventory, course_enrollments, and student_details. Those are relational data since they have relations so that they can connect with others via one or several variables. Before joining those tables into one data frame, I explore those tables first.

### Load data


```{r}
acad_plan <- read.csv(file = "D:/R/data analysis/Institutional research/technical-exercise/content/docs/data/acad_plan.csv")
class_instructors <- read.csv(file = "D:/R/data analysis/Institutional research/technical-exercise/content/docs/data/class_instructors.csv")
class_inventory <- read.csv(file = "D:/R/data analysis/Institutional research/technical-exercise/content/docs/data/class_inventory.csv")
course_enrollments <- read.csv(file = "D:/R/data analysis/Institutional research/technical-exercise/content/docs/data/course_enrollments.csv")
student_details <- read.csv(file = "D:/R/data analysis/Institutional research/technical-exercise/content/docs/data/student_details.csv")
```

+  **acad_plan** describes detail information related to academic plan
```{r echo=FALSE}
head(acad_plan,6)
```

+ **class_instructors** contains instructor id, class number, and term code
```{r echo=FALSE}
head(class_instructors,6)
```

+ **class_inventory** includes detail information relate to class, such as term code, class number, etc.
```{r echo=FALSE}
head(class_inventory,6)
```

+ **course_enrollments** indicates student performance and course details
```{r echo=FALSE}
head(course_enrollments,6)
summary(course_enrollments)
```
hs_gpa_entry variable has several extreme values, such as 533. I drop those observations.
```{r}
student_details <- student_details %>%
  mutate(
    hs_gpa_entry = case_when(
      hs_gpa_entry <= 5 ~hs_gpa_entry, 
      TRUE ~ NA
    )
  )
```


+ **student_details** represents characteristics of students
```{r echo=FALSE}
head(student_details,6)
```

```{r eval=FALSE, include=FALSE}
class_instructors %>%
  count(class_nbr, term_code) %>%
  filter(n>1) %>%
  nrow()
```

```{r eval=FALSE, include=FALSE}
class_inventory%>%
  count(class_nbr, term_code) %>%
  filter(n>1) %>%
  nrow()
```

```{r eval=FALSE, include=FALSE}
course_enrollments%>%
  count(class_nbr, term_code, student_id_number) %>%
  filter(n>1) %>%
  nrow()
```

```{r eval=FALSE, include=FALSE}
acad_plan%>%
  count(acad_plan, dept_code,degree,college,eff_status, acad_plan_desc) %>%
  filter(n>1) %>%
  nrow()
```

```{r include=FALSE}
df <- course_enrollments %>% 
  full_join(student_details, c("student_id_number")) %>%
  full_join(class_instructors, c("term_code", "class_nbr")) %>%
  full_join(class_inventory,c("term_code","class_nbr")) %>%
  full_join(acad_plan, c("acad_plan", "dept_code")) 
```

```{r include=FALSE}
dt <- format(Sys.time(),"%y%m%d_%H%M")
fileName <- paste0("data_", dt, ".Rdata",sep="")
fileName
save.image(fileName)
```

